---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rconnect

<!-- badges: start -->
<!-- badges: end -->

rconnect is a simple package that implements riverscape connectivity measures for wind dispersed plants as 
described by Wagner & Wöllner (2022):

*absoluteConnections*  
*effectiveConnections*  
*effectiveDistance*  
*effectiveSeedrain*  

It further provides s helper function to create a negative exponential dispersal kernel and simple example raster file with suitable habitats.

## Installation

You can install the current development version of rconnect from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TCWagner/rconnect")
```

## Example

This is a basic example how to work with the package and calculate simple connectivity metrics of a riverscape.

Let's find out how well the habitats of a riverscape (a riverine landscape) are connected for a certain species. As example we chose the Asteracea *Chondrilla chondrilloides*,
an wind dispersed species well adapted to open gravel bars of alpine rivers. For details about the species ecology, dispersal and status, see Wöllner et al. 2022.

We start with a raster file, containing the suitable habitats for our species. Suitable habitats need to have a value > 0; 0 codes for unsuitable habitats. Here we use our example data that comes with our package: *habitats_lech*

```{r example}
library(raster)
library(rconnect)

## load the example raster with suitable habitats. suitable habitats need to be coded with values > 0
## you may use your own raster, here we use our example data:
data(habitats_lech)
plot(habitats_lech)
```

Now we the dispersal kernel for the species under consideration as a matrix. We can create a simple, negative exponential dispersal kernel with the function *dispersalKernel*. 
Assuming a negative exponential decrease of the seeds with distance and a dispersal distance of ~14m the decay is 0.19. 

To create the kernel, we need to provide the cell size of our raster containing the suitable habitats (5m), and the intended radius of our kernel in cells. By default, the kernel center cell will be set to 0 and the kernel normalized to sum up to 1.

```{r kernel}

cckernel <- dispersalKernel(cellsize=5, radius=5, decay=0.19)

```

Now we can easily calculate the number of connections (eC) that each patch has:

```{r nC}

nC <- absoluteConnections(habitats_lech, cckernel)
nC

```

However, if we want to have the effective connections or *eC* (that is the connections weighted by distance) we can use:

```{r eC}

eC <- effectiveConnections(habitats_lech, cckernel)
eC

```


If we want to have the effective connectivity for our whole riverscape we just need to calculate the mean of the effective connections:

```{r eCC}

mean(eC$eC)

```
So, our riverscape has a conectivity of ~30% for Chondrilla chondrilloides.

We can now determine the effective distance (*eD*), a measure that tells us, how fas away a patch would be from an ... 

```{r eD}

eD <- effectiveDistance(habitats_lech, cckernel)
eD
```

You may wish to have your results a a raster file for further analysis instead of a simple table. No problem, just call the respective functions with *summarize=FALSE* option and you will get a raster where each patch is assigned the respective value:

```{r eDr}

eDr <- effectiveDistance(habitats_lech, cckernel, summarize=FALSE)
plot(eDr)
```

Sometimes you may wish to have the cell-by-cell connectivity, or *effectiveSeedrain* to use for further modeling:

```{r eCr}

eCr <- effectiveSeedrain(habitats_lech, cckernel, summarize=FALSE)
plot(eCr)
```

And, finally, you may wish to have the effective patch-to-patch connectivity and the respective effective distances between patches

```{r eCM}

eCM <- effectiveConnectionsMatrix(habitats_lech, cckernel)
eCM
```

Please note, that the effective distances for a patch cannot be summed up or averaged to obtain its total *effectiveDistance*. Use the appropriate function *effectiveDistance* instead. 

## Finally


The basic functions of this package, *effectiveConnections* and *effectiveSeedrain*  provide spatially explicit data if needed. Though currently we do neither consider the actual occupancy of habitats or barriers, both can easily be combined with our functions. The respective data can be used for further modelling. We do not include long distance dispersal here, because here the mechanisms are different and more complex. AA package suitable for this is currently under development.

## References

Woellner, R., Bräuchler, C., Kollmann, J., & Wagner, T. C. (2022). Biological Flora of Central Europe: Chondrilla chondrilloides (Ard.) H. Karst. Perspectives in Plant Ecology, Evolution and Systematics, 54, 125657.

Wagner, T. C., Woellner, R. (2022). Effective Connectivity and effective Habitat Distance - A new metric to quantify habitat connectivity for plant species in riverscapes. Ecological Indicators.

