---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rconnect

<!-- badges: start -->
<!-- badges: end -->

rconnect is a simple package that implements riverscape connectivity
measures for quantifying the colonization potential of wind dispersed plants species as described by Wagner & Wöllner
(2023):

*effectiveSeedrain*  
*effectiveConnections*  
*effectiveDistance*  
*colonizationPotential*
*absoluteConnections*
*connectionCapacity*

It further provides a helper function to create a negative exponential dispersal kernel, functions to create a connectivity and distance matrix based on effectiveConnections and effectiveDistance and a simple example raster file with suitable habitats.

## Installation

You can install the current development version of rconnect from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TCWagner/rconnect")
```

## Example

Here is a basic example how to work with the package and calculate simple connectivity metrics of a riverscape.

Let's find out how well the habitats of a riverscape (a riverine landscape) are connected for a certain species. As example we chose the Asteracea *Chondrilla chondrilloides*,
an wind dispersed species well adapted to open gravel bars of alpine rivers. For details about the species ecology, dispersal and status, see Wöllner et al. 2022.

We start with a raster file, containing the suitable habitats for our species. Suitable habitats need to have a value > 0; 0 codes for unsuitable habitats. Here we use our example data that comes with our package: *habitats_lech*

```{r example}
library(raster)
library(rconnect)

## load the example raster with suitable habitats. suitable habitats need to be coded with values > 0
## you may use your own raster, here we use our example data:
data(habitats_lech)
plot(habitats_lech)
```

Now we need the dispersal kernel for the species under consideration as a matrix. We can create a simple, negative exponential dispersal kernel with the function *dispersalKernel*. 
Assuming a negative exponential decrease of the seeds with distance and a dispersal distance of ~14m the decay is 0.19. 

To create the kernel, we need to provide the cell size of our raster containing the suitable habitats (5m), and the intended radius of our kernel in cells. By default, the kernel center cell will be set to 0 and the kernel normalized to sum up to 1.

```{r kernel}

cckernel <- dispersalKernel(cellsize=5, radius=7, decay=0.19)

```

Lets start to calculate the seed rain that can be exchanged between neighboring habitats (*effectiveSeedrain*). If we do not *summarize*, the output will be a raster with the relative contribution that each cell of an habitat has.

```{r eS}

eS <- effectiveSeedrain(habitats_lech, cckernel, summarize=F)
plot(eS)

```

Now, lets see how many *effective Connections* a patch has with the other patches of the riverscape

```{r eCM}

eCM <- effectiveConnectionsMatrix(habitats_lech, cckernel)
eCM
```




However, if we want to have the total *effective Connections* or *eC* that a patch has with all neighbors (that is the connections weighted by distance) we can use:

```{r eC}

eC <- effectiveConnections(habitats_lech, cckernel)
eC

```

We may want to have this a a raster file that we can use for further modeling. So lets set *summarize* to *FALSE*:

```{r eCr}

eCr <- effectiveConnections(habitats_lech, cckernel, summarize=F)
plot(eCr)

```

Similarly, we can calculate the *effective Distance* for each patch (*eD*):

```{r eD}

eD <- effectiveDistance(habitats_lech, cckernel)
eD
```

We may wish to have your results a raster file for further analysis instead of a simple table. No problem, we just call the respective functions with *summarize=FALSE* option and  will get a raster where each patch is assigned the respective value:

```{r eDr}

eDr <- effectiveDistance(habitats_lech, cckernel, summarize=FALSE)
plot(eDr)
```


Finally we want a summary of our metrics and the total *colonization Potential* of our riverscape:


```{r eCP}

cP <- colonizationPotential(habitats_lech, cckernel)
cP

```

So, our riverscape has a *colonization Potential* of 0.776 (~78%) for *Chondrilla chondrilloides*. The average *effective Distance* a patch has is 7.7m, clearly within the species dispersal distance. On average, each patch has 2 connections with a capacity of 0.33. However, the relative standard deviation of all parameters is quite high, indicating an unequal distribution of the connectivity. Note, that the average *effective Distance* (*eDm*) does only consider connected patches. *nCm* gives us the average *number of Connections* a patch has, and *cCm* tells us the average *connection Capacity* (relative amount of seeds) that can be exchanged by a connection.








## Final Note


The basic functions of this package, *effectiveConnections* and *effectiveSeedrain*  provide spatially explicit data if needed. Though currently we do neither consider the actual occupancy of habitats or barriers, both can easily be combined with our functions. The respective data can be used for further modelling. We do not include long distance dispersal here, because here the mechanisms are different and more complex. AA package suitable for this is currently under development.

## References

Woellner, R., Bräuchler, C., Kollmann, J., & Wagner, T. C. (2022). Biological Flora of Central Europe: Chondrilla chondrilloides (Ard.) H. Karst. Perspectives in Plant Ecology, Evolution and Systematics, 54, 125657.

Wagner, T. C., Woellner, R. (2022). Effective Connectivity and effective Habitat Distance - A new metric to quantify habitat connectivity for plant species in riverscapes. Ecological Indicators.

